---
title: "CDSimX-introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CDSimX-introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  fig.height = 4,
  comment = "#>"
)
```

```{r setup}
library(CDSimX)
```

<h2 style="font-size:160%; font-weight:bold;"> Introduction </h2>

The CDSimX package provides an advanced framework for climate simulation,
forecasting, visualization, validation, and climate data export in R.

The package supports:

- synthetic weather station generation

- stochastic climate simulation

- machine learning forecasting

- climate validation

- NetCDF and CSV export

- visualization of climate variables

- climate dependence modelling

CDSimX is useful for:

- climate research

- hydrological modelling

- machine learning experiments

- educational demonstrations

- sensitivity analysis

- simulation studies

<h2 style="font-size:160%; font-weight:bold;"> Creating Weather Stations </h2>
The stations can be created either by: loading from a CSV file, accepting an existing data frame, or auto-generating synthetic stations in a bounding box. We begin by generating synthetic climate stations.
```{r}
stations <- create_stations(
  n = 3,
  seed = 123
)
stations
```
Each station contains:

- station identifier

- longitude

- latitude

- elevation (if available)

<h2 style="font-size:160%; font-weight:bold;"> Generating Time Indices </h2>

CDSimX supports daily, monthly, and yearly temporal resolutions.
```{r}
time_index <- generate_time_index(
  start_date = "2019-01-01",
  end_date = "2024-12-31",
  frequency = "day"
)
head(time_index)
```
<h2 style="font-size:160%; font-weight:bold;"> Simulating Climate Data </h2>

Climate variables can now be simulated using the generated stations and
time index.
```{r}
climate <- simulate_climate(
  stations = stations,
  time_index = time_index,
  seed = 123
)

head(climate)
```
The simulated dataset may include:

- Tmin

- Tmax

- Rainfall

- RH

- WindSpeed

- Solar_Radiation

- ET0

- DewPoint

<h2 style="font-size:160%; font-weight:bold;"> Climate Visualization </h2>

CDSimX includes customizable visualization functions.
```{r}
plot_station_timeseries(
  climate,
  station = "Station_1",
  var = "Tmax"
)
```
Users may customize:

- colors

- themes

- smoothing

- labels

- seasonal highlighting

<h2 style="font-size:160%; font-weight:bold;"> Machine Learning Forecasting </h2>

CDSimX supports several machine learning forecasting approaches.

Available methods include:

- Random Forest

- Linear Regression

- GBM

- ARIMA

- XGBoost

- Neural Networks

```{r}
forecast_result <- forecasting_ml(
  climate_data = climate,
  target = "Rainfall",
  forecast_horizon = 12,
  method = "rf"
)
forecast_result$model_performance
```
Forecasted values:
```{r}
head(forecast_result$forecast_data)
```

<h2 style="font-size:160%; font-weight:bold;"> Climate Validation </h2>

CDSimX includes climate validation tools for checking realism and
statistical consistency.
```{r}
validation <- validate_climate(climate)
validation
```
Validation diagnostics may include:

- physical constraint checks

- variability assessment

- correlations

- temporal consistency

- distributional diagnostics

<h2 style="font-size:160%; font-weight:bold;"> Exporting Climate Data </h2>

<h4 style="font-size:160%; font-weight:bold;"> Export to CSV </h4>

```{r}
export_csv(
  climate,
  file = "simulated_climate.csv"
)
```

<h4 style="font-size:160%; font-weight:bold;"> Export to NetCDF </h4>
```{r}
export_netcdf(
  climate,
  file = "simulated_climate.nc"
)
```

These formats support interoperability with:

- ncdf4

- terra

- stars

- xarray

- climate modelling workflows

<h2 style="font-size:160%; font-weight:bold;"> Conclusion </h2>

CDSimX provides a modern climate simulation ecosystem for generating,
forecasting, validating, visualizing, and exporting synthetic climate data.

The package is designed to support reproducible climate science,
machine learning applications, and environmental modelling workflows.
